Embedding Dimension of Polyhedral Losses

### 2020-01-27 colt-20/poly-embed.tex
Filling in some TO DO items

The main result of this section is Theorem \ref{thm:orderable-1-embed}.
\begin{theorem} \label{thm:orderable-1-embed}
  Given a discrete loss $\ell: \R \to \nonnegreals^{\Y}$, the following are equivalent:
  \begin{enumerate}
    \item The property elicited by $\ell$ is orderable.
    \item $\ell$ is $1$-embeddable.
    \item $\ell$ has some polyhedral calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
    \item $\ell$ has some convex calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
  \end{enumerate}
\end{theorem}
This theorem implies that one can directly test if a property is $1$-embeddable via the orderable criterion.
Orderability can be tested using the Bayes risk of $\ell$, a polyhedral function defined by $p \mapsto \max_{r \in \R} \inprod{p}{\ell(r)}$.
A direct approach is to project this function down to the simplex to give the level sets $\gamma_r$ \bo{figure?}.
Another approach is to note that the intersection graph is directly given by finding those pairs $r,r'$ that simultaneously maximize the Bayes risk, i.e. cases where for some $p$, we have $r,r' \in \argmax_{r'' \in \R} \inprod{p}{\ell(r'')}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Higher dimensions}
\bo{blurb about the purpose of this section and the layout.}

\subsection{Setup: subgradient sets at embedding points.}
%Recall that we have a loss function $\ell: \R \to \reals^{\Y}_{\geq 0}$, and for any prediction $r \in \R$ of the algorithm, we write $\gamma_r = \{ p : \inprod{p}{\ell(r)} \leq \inprod{p}{\ell(r')} (\forall r')\}$.
%We are interested in potential polyhedral surrogates $L: \reals^d \to \reals^{\Y}_{\geq 0}$.
%These are associated with embedding points, where $u_r \in \reals^d$ is linked back to the discrete prediction $r$.

@@

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} p \cdot \ell(r)\}$.
  Then $\ell$ is $d$-embeddable if and only if there exist polytopes $\T = \{T^r_y \subseteq \reals^d : r \in \R, y \in \Y\}$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) For all $r \in \R$, we have $D(\T^r) = \gamma_r$.
    \item (Monotonicity) (TODO)
  \end{enumerate}
\end{theorem}
\begin{proof}
  TODO
\end{proof}

\subsection{Characterizing optimality}

@@

\begin{corollary}
  If a discrete loss $\ell$ has any level set $C = \gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r')} \}$ for which the quadratic program is infeasible with parameter $d$, then $\ell$ is not $d$-embeddable.
\end{corollary}

>>>>

The main result of this section is Theorem \ref{thm:orderable-1-embed}.
\begin{theorem} \label{thm:orderable-1-embed}
  Given a discrete loss $\ell: \R \to \nonnegreals^{\Y}$, the following are equivalent:
  \begin{enumerate}
    \item The property elicited by $\ell$ is orderable.
    \item $\ell$ is $1$-embeddable.
    \item $\ell$ has some polyhedral calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
    \item $\ell$ has some convex calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
  \end{enumerate}
\end{theorem}
This theorem implies that one can directly test if a property is $1$-embeddable via the orderable criterion.
Orderability can be tested using the Bayes risk of $\ell$, a polyhedral function defined by $p \mapsto \max_{r \in \R} \inprod{p}{\ell(r)}$.
A direct approach is to project this function down to the simplex to give the level sets $\gamma_r$, as in Figure~\ref{fig:intersection-graph}.
Another approach is to note that the intersection graph is directly given by finding those pairs $r,r'$ that simultaneously maximize the Bayes risk, i.e. cases where for some $p$, we have $r,r' \in \argmax_{r'' \in \R} \inprod{p}{\ell(r'')}$.

\begin{figure}\label{fig:intersection-graph}
	\centering
	\includegraphics[width = 0.6\linewidth]{tikz/intersection-graph.pdf}
	\caption{Intersection graph for the [truncated] expected value property on outcomes $\Y = \{1,2,3\}$.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Higher dimensions}
\bo{blurb about the purpose of this section and the layout.}
\jessie{Taking a crack at it:}
Our newfound understanding of the 1-dimensional embedding characterization leads us to realize that a large class of properties are not $1$-embeddable.
\cite{finocchiaro2019embedding} have shown a trivial upper bound that every property is $(n-1)$-embeddable, but it is an open question as to how tight we can get the bounds on embedding dimension.
Reducing embedding dimension significantly below $(n-1)$ can help provide a computational speedup in the optimization problem, which is a function of the surrogate loss dimension.
This is especially important as the number of possible outcomes $n$ grows large.

Below, we generalize our characterization to $d$-embeddability for $d \geq 2$.
Having a finite number of embedding points and restricting to polyhedral functions allows us to study the subgradient sets of the loss at embedding points, which we formalize in Section~\ref{subsec:sub-sets}.
Using a general theory of polytopes, we can then present our general $d$-dimensional characterization in terms of optimality and monotonicity conditions.
In Section~\ref{subsec:opt-conditions}, we present a quadratic feasibility program in order to understand necessary conditions for $d$-embeddability by satisfying optimality conditions, yielding new lower bounds for embedding dimension in Corollary~\ref{cor:d-embeddable-char}.

\subsection{Setup: subgradient sets at embedding points.}\label{subsec:sub-sets}
%Recall that we have a loss function $\ell: \R \to \reals^{\Y}_{\geq 0}$, and for any prediction $r \in \R$ of the algorithm, we write $\gamma_r = \{ p : \inprod{p}{\ell(r)} \leq \inprod{p}{\ell(r')} (\forall r')\}$.
%We are interested in potential polyhedral surrogates $L: \reals^d \to \reals^{\Y}_{\geq 0}$.
%These are associated with embedding points, where $u_r \in \reals^d$ is linked back to the discrete prediction $r$.

@@

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} p \cdot \ell(r)\}$.
  Then $\ell$ is $d$-embeddable if and only if there exist polytopes $\T = \{T^r_y \subseteq \reals^d : r \in \R, y \in \Y\}$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) For all $r \in \R$, we have $D(\T^r) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\phi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $y \in \Y$ and $r \in \R$, we have $T^r_y = \partial L_y(\phi(r))$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  TODO
\end{proof}

\subsection{Characterizing optimality}\label{subsec:opt-conditions}

@@

\begin{corollary}\label{cor:d-embeddable-char}
  If a discrete loss $\ell$ has any level set $C = \gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r')} \}$ for which the quadratic program is infeasible with parameter $d$, then $\ell$ is not $d$-embeddable.
\end{corollary}
